\documentclass{article}

\usepackage[margin=1in]{geometry}

\usepackage{amsfonts}
\def\R{{\mathbb R}}
\def\N{{\mathbb N}}

\title{In-Class Exercise 04}
\author{CSCI 432}

\begin{document}
\maketitle

\noindent
Group Number:\\
Group members present today:

\section*{Minimum Enclosing Disc (MED)}

\paragraph{Definitions:}

\begin{enumerate}
\item Given $c \in \R^2$ and $r \in \R$ such that $r \geq 0$, we define the
disc $D(c,r) := \{ x \in \R^2 ~|~ ||x-c|| \leq r \}$.
\begin{itemize}
\item Draw $D(0,1)$ and $D(0,2)$. Note: $0\in \R^2$ is the origin
$(0,0)$.
\vspace{3ex}
\item Let $d>1$.  The generalization of a two-dimensional disc is a
the \emph{Euclidean metric ball}.  Give a general definition of
a ball in $\R^d$.  Denote this ball by $\mathbb{B}_d(c,r)$.
\vspace{3ex}
\item Draw $\mathbb{B}_1(0,1)$.
\vspace{3ex}
\end{itemize}
\item A \emph{circle} is the boundary of the disc.  What is an equation that
defines the circle $C(c,r)$?
\vspace{3ex}
\end{enumerate}

\paragraph{Problem Statement:}

Let $P \subset \R^2$, with $|P|=n \in \N$.  We wish to find the smallest radius
$r$ such that there exists a $c \in \R^2$, where $P \subset D(c,r)$.

\begin{enumerate}
\item If $n=1$, what is the minimum enclosing disc?  Is it unique?
\vspace{3ex}
\item If $n=2$, what is the minimum enclosing disc?  Is it unique?
\vspace{3ex}
\item If $n=3$, what is the minimum enclosing disc?  Is it unique?
\vspace{3ex}
\item If $n=4$, what are the possible cases that could arise?  How do we
decide what the MED is?
\vspace{3ex}
\item Use the following to consider the general case:
consider the following: choose a point $p$ at random.  Remove $p$ from $P$
to obtain $P'$
and compute SEB of $P'$.  What are the two cases that can happen when we
add $p$ back in?  What is the probability of each?
\vspace{3ex}
\item For the expected time analysis, what is the recursion that we have?
What is the closed form?
\vspace{3ex}
\item Challenge: In $\R^d$, how many points are needed in order to uniquely define a
ball whose boundary contains those points?
\vspace{3ex}
\end{enumerate}

\end{document}